% Exam Template for UMTYMP and Math Department courses
%
% Using Philip Hirschhorn's exam.cls: http://www-math.mit.edu/~psh/#ExamCls
%
% run pdflatex on a finished exam at least three times to do the grading table on front page.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% These lines can probably stay unchanged, although you can remove the last
% two packages if you're not making pictures with tikz.
\documentclass[11pt]{exam}
\RequirePackage{amssymb, amsfonts, amsmath, latexsym, verbatim, xspace, setspace}


% By default LaTeX uses large margins.  This doesn't work well on exams; problems
% end up in the "middle" of the page, reducing the amount of space for students
% to work on them.
\usepackage[margin=1in]{geometry}
\usepackage[english]{babel}
\usepackage[autostyle]{csquotes} %%%% This package allows Tex to recognize quotation marks with the \enquote command. 


% Here's where you edit the Class, Exam, Date, etc.
\newcommand{\class}{MAT 203}
\newcommand{\term}{Summer II 2018}
\newcommand{\examnum}{Midterm}
\newcommand{\examdate}{07/26/18}
\newcommand{\timelimit}{3 hours and 25 minutes}

% For an exam, single spacing is most appropriate
\singlespacing
% \onehalfspacing
% \doublespacing

% For an exam, we generally want to turn off paragraph indentation
\parindent 0ex
\title{MAT 203 - Summer II 2018: Midterm}
\begin{document} 

% These commands set up the running header on the top of the exam pages
\pagestyle{head}
\firstpageheader{}{}{}\textbf{}
\runningheader{\class}{\examnum\ - Page \thepage\ of \numpages}{\examdate}
\runningheadrule

\begin{flushright}
\begin{tabular}{p{2.8in} r l}
\textbf{\class} & \textbf{Name (Print):} & \makebox[2in]{\hrulefill}\\
\textbf{\term} &&\\
\textbf{\examnum} &&\\
\textbf{\examdate} &&\\
\textbf{Time Limit: \timelimit} & ID number & \makebox[2in]{\hrulefill}
\end{tabular}\\
\end{flushright}
\rule[1ex]{\textwidth}{.1pt}

\begin{center}
\large{\textbf{Instructions}}
\end{center}

\begin{minipage}[t]{3.7in}
\vspace{0pt}
\begin{itemize}

\item This exam contains \numpages\ pages (including this cover page) and
\numquestions\ problems.  Check to see if any pages are missing.  Enter
all requested information on the top of this page, and put your initials
on the top of every page, in case the pages become separated.

\item You may \textit{not} use your books, notes, or any device that is capable of accessing the internet on this exam (e.g., smartphones, smartwatches, tablets). You may not use a calculator.

\item \textbf{Organize your work}, in a reasonably neat and coherent way, in
the space provided. Work scattered all over the page without a clear ordering will 
receive very little credit.  

\item \textbf{Mysterious or unsupported answers will not receive full
credit}.

\end{itemize}

\end{minipage}
\hfill
\begin{minipage}[t]{2.3in}
\vspace{0pt}
%\cellwidth{3em}
\gradetablestretch{2}
\vqword{Problem}
\addpoints % required here by exam.cls, even though questions haven't started yet.	
\gradetable[v]%[pages]  % Use [pages] to have grading table by page instead of question

\end{minipage}
\newpage % End of cover page

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% See http://www-math.mit.edu/~psh/#ExamCls for full documentation, but the questions
% below give an idea of how to write questions [with parts] and have the points
% tracked automatically on the cover page.
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{questions}

% Basic question
%%%%%%%%%%%%%%

\addpoints
\question In each of the following problems, determine if the statements are true or false. Explain your reasoning (correct answers without an explanation will be worth only 2 points per statement). 
\begin{parts}
\part[5] Two vectors in space always determine a unique plane. 
\vfill
\part[5] Two planes in space always intersect. 
\vfill
\part[5] The dot product can be used to detect whether two non-zero vectors in space are aligned. 
\vfill
\part[5] Let $r(t)$ and $s(t)$ be curves in the plane, such that neither has a limit as $t$ converges to $0$. Then their cross product $r(t)\times s(t)$ does not have a limit at $0$ either. 
\vfill
\part[5] If all directed limits a scalar-valued, multivariable function at a point exist and coincide, then the function has a limit at the point, in the multivariable sense.
\vfill
\part[5] If a scalar-valued, multivariable function is separately continuous with respect to each variable, then it is continuous in the multivariable sense.
\vfill
\end{parts}

\newpage 


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\addpoints 
\question Consider the lines whose parametric equations are given by 
\begin{align*}
L_1 \colon  & x=2t, y=4t, z=6t.\\
L_2 \colon & x=1-s, 4+s, -1+s.
\end{align*}

\begin{parts}
\part[5] Explain why these lines do not intersect. 
\vfill
\part[5] Explain why these lines are not parallel. 
\vfill
\newpage
\part[5]  Find the general equations of two parallel planes, $\Pi_1$ and $\Pi_2$, containing lines $L_1$ and $L_2$, respectively. 
\vfill
\part[5] What is the distance between the lines $L_1$ and $L_2$ (this is the same as the distance between the planes $\Pi_1$ and $\Pi_2$ above).
\vfill
\end{parts}

\newpage

%%%%%%%%%%%%%%%%%
\addpoints
\question[10] Compute the trajectory of a curve whose velocity vector is given by 
\begin{equation*}
r'(t)= e^{-t}i + t^2j + \frac{1}{1+t^2} k,
\end{equation*}
and such that $r(0)=(1,1,1)$. 
\newpage
%%%%%%%%%%%%%%%%%

\addpoints

\addpoints
\question Consider the function
\begin{equation*}
f(x,y)=\left\{
\begin{array}{rc}
\frac{xy(x^2-y^2)}{x^2+y^2} & \mbox{if} \ (x,y) \neq (0,0)\\
0, & \mbox{if} \ (x,y)=(0,0).
\end{array}
\right.
\end{equation*}

\begin{parts}
\part[5] Where are the directed limits of the function along lines $y=kx$, as $x \to 0$?
\vfill
\part[5] Is this function continuous at the origin?
\vfill
\newpage
\part[5] What is the partial derivative of this function relative to $x$, at points other than the origin?
\vfill
\part[5] What is the partial derivative of this function relative to $y$, at points other than the origin?
\vfill
\end{parts}
\newpage

%%%%%%%%%%%%%%%%%%%%
\addpoints
\question Consider a model of gravitation in which two bodies, A and B, exert gravitational pull on one another. Assume that the mass of A is much bigger than that of $B$, so that the acceleration felt by $A$ is negligible, and we can regard it as in rest (this is a good approximation for the orbital motion of artificial satelites around the Earth, for instance). We are going to consider our reference frame for 3-space as centered at $A$, and describe the position of $B$ by a vector-valued function of time, $r(t)$. 

Newton's Law of Universal Gravitation asserts that the gravitational force exerted by $A$ upon $B$ is of the form
\begin{equation*}
F=\frac{K}{||r(t)||^3} r(t),
\end{equation*}
where $K$ is a (negative) constant depending on the masses of $A$ and $B$. Meanwhile, Newton's Second Law of Motion says that the acceleration $r''(t)$ felt by $B$ as a result of the gravitational pull from $A$ can be computed by the well-known formula
\begin{equation*}
F=m r''(t),
\end{equation*}
where $m$ is the mass of $B$ (a constant). 

\begin{parts}
\part[10] By comparing the two expressions for the force, explain why the cross product of position and velocity vectors, $r(t) \times r'(t)$, is constant. 
\vfill 
\newpage 
\part[10] Use part (a) to explain why the position vector $r(t)$ belongs to a fixed plane, for all $t$.  
\vfill
\end{parts}

%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{questions}
\end{document}
